# Initial Data Analysis
*This is the first of a six-part series created to analyze and predict blight ticket compliance in the city of Detroit. You can find links to the other articles in this series at the bottom of the post.* 

The primary concern of the initial data analysis is data quality and engages questions such as:  
1. How should missing and faulty data be treated?   
2. Do data properties agree with expectations?  
3. What initial data transformations would be appropriate?  
4. What are the distribution and frequency characteristics of categorical variables?    
5. What inconsistencies exist and what should be done about them?  
6. What outliers exist and what are the implications for further statistical analyses?  

This article will examine the Detroit Blight Ticket Data in some detail, examine the data and their properties, illuminate any inconsistencies and prescribe treatments and corrective actions which will be implemented in the following data preparation phase.

## Blight Ticket Data
The data for this project was made available courtesy of the publicly accessible Detroit Open Data Portal. Available blight ticket data has been split into a training set containing observations from 2005-2011 and a test set including blight tickets written between 2012 and 2016. Each row in these two files corresponds to a single blight ticket, and includes information about when, why, and to whom each ticket was issued. The target variable is compliance, which is True if the ticket was paid early, on time, or within one month of the hearing date, False if the ticket was paid after the hearing date or not at all, and Null if the violator was found not responsible. Compliance, as well as a handful of other variables that will not be available at test-time, are only included in train.csv. In addition, all tickets where the violators were determined to be "not responsible" are not considered during evaluation. They are included in the training set as an additional source of data for visualization, and to enable unsupervised and semi-supervised approaches. Here are the variables and their definitions.

train.csv & test.csv

* ticket_id - unique identifier for tickets   
* agency_name - Agency that issued the ticket  
* inspector_name - Name of inspector that issued the ticket  
* violator_name - Name of the person/organization that the ticket was issued to   
* violation_street_number, violation_street_name, violation_zip_code - Address where the violation occurred  
* mailing_address_str_number, mailing_address_str_name, city, state, zip_code, non_us_str_code, country - Mailing address of the violator  
* ticket_issued_date - Date and time the ticket was issued   
* hearing_date - Date and time the violator's hearing was scheduled   
* violation_code, violation_description - Type of violation   
* disposition - Judgment and judgement type   
* fine_amount - Violation fine amount, excluding fees   
* admin_fee - $20 fee assigned to responsible judgments   
* state_fee - $10 fee assigned to responsible judgments   
* late_fee - 10% fee assigned to responsible judgments   
* discount_amount - discount applied, if any   
* clean_up_cost - DPW clean-up or graffiti removal cost   
* judgment_amount - Sum of all fines and fees   
* grafitti_status - Flag for graffiti violations   

train.csv only    

* payment_amount - Amount paid, if any  
* payment_date - Date payment was made, if it was received   
* payment_status - Current payment status as of Feb 1 2017   
* balance_due - Fines and fees still owed  
* collection_status - Flag for payments in collections   
* compliance [target variable for prediction]   
+ Null = Not responsible  
 0 = Responsible, non-compliant  
 1 = Responsible, compliant  
* compliance_detail - More information on why each ticket was marked compliant or non-compliant  

```{python ida_read, code = readLines("../ida.py")[2:34]}
```

The blight dataframe summarized contains `r nrow(py$df)` blight ticket observations and `r ncol(py$df)` variables.

Two additional files, addresses.csv and latlons.csv have been provided for geolocation purposes. The former contains the violation addresses mapped to ticket_ids and the latter maps the addresses to latitude and longitude coordinates. 

Not every variable will be relevant to, or useful for, analysis and modeling. Let's trim the dataset down to the relevant observations and variables.

* * *  
## Data Selection
Here, we will identify the relevant observations and variables and eliminate the superfluous, redundant, and tangential data. We are only interested in tickets in which the violator was deemed 'responsible', those tickets with a 'not responsible' designation will be removed. Variables containing payment and balance information are only included in the training set and can therefore be discarded. Redundant variables such as those containing fees, costs and discounts can also be eliminated from further analysis. 

```{python ida_select, code = readLines("../ida.py")[39:46]}
```

The trimmed dataframe described in `r kfigr::figr(label = "ida_vars", prefix = TRUE, link = TRUE, type="Figure")` now contains `r nrow(py$df)`  observations and the following `r ncol(py$df)` variables.  We can now proceed with the univariate data analysis. 

```{python ida_vars, echo=FALSE}
print(df.info())
```
`r kfigr::figr(label = "ida_vars", prefix = TRUE, link = TRUE, type="Figure")`: Selected Blight Ticket Data Variables

* * *  

## Univariate Analysis
Again, the univariate data analysis is primarily concerned with characterizing the quality of the data. The following descriptive statistics and frequency analyses will inform our downstream data processing, feature selection, feature engineering and modeling decisions. 

Let's start with the target variable, Compliance. Por qué no? 

### Compliance  
```{python compliance, code = readLines("../ida.py")[53:59]}
```
The compliance variable contains either zero or one for all observations as expected. As graphically depicted in `r kfigr::figr(label = "compliance_plot", prefix = TRUE, link = TRUE, type="Figure")`, the vast majority of blight tickets (`r round(py$compliance$Percent[2], 1)`%) were 'Non-Compliant' and there are no missing values.

```{python compliance_plot, code = readLines("../ida.py")[61:63]}
```
`r kfigr::figr(label = "compliance_plot", prefix = TRUE, link = TRUE, type="Figure")`: Compliance Summary

`r kfigr::figr(label = "compliance_tbl", prefix = TRUE, link = TRUE, type="Table")`: Compliance Distribution
```{r compliance_tbl}
knitr::kable(py$compliance, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Though no data quality issues present at this stage, a `r round(py$compliance$Percent[1], 1)`% compliance rate is rather abysmal indeed.

* * * 
### Agency
```{python agency, code = readLines("../ida.py")[70:73]}
```
`r kfigr::figr(label = "agency_plot", prefix = TRUE, link = TRUE, type="Figure")` shows `r sum(py$agency$Count)` blight tickets originating from one of the five enforcement agencies and there are no missing values for this variable.

```{python agency_plot, code = readLines("../ida.py")[75:78]}
```
`r kfigr::figr(label = "agency_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Tickets by Agency

`r kfigr::figr(label = "agency_tbl", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Agency 
```{r agency_tbl}
knitr::kable(py$agency, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As shown in `r kfigr::figr(label = "agency_tbl", prefix = TRUE, link = TRUE, type="Table")`, Neighborhood City Halls are responsible for just a single blight ticket. To ensure an adequate number of observations per categorical level, we'll combine the Neighhborhood City Halls and Detroit Police Department levels into a single categorical level.

* * *
### Inpector
```{python inspector, code = readLines("../ida.py")[85:94]}
```
`r kfigr::figr(label = "inspector_plot", prefix = TRUE, link = TRUE, type="Figure")` shows `r sum(py$inspector$Count)` blight tickets originating from one of the `r py$inspector_summary$unique` inspectors and there are no missing values for this variable.

`r kfigr::figr(label = "inspector_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Inspector Summary
```{r inspector_summary}
knitr::kable(py$inspector_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python inspector_plot, code = readLines("../ida.py")[95:98]}
```
`r kfigr::figr(label = "inspector_plot", prefix = TRUE, link = TRUE, type="Figure")`: Inspector Blight Ticket Frequency Histogram

`r kfigr::figr(label = "inspector_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Inspector Frequency Analysis
```{r inspector_spectrum}
knitr::kable(py$inspector_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As indicated in `r kfigr::figr(label = "inspector_plot", prefix = TRUE, link = TRUE, type="Figure")`, the distribution of blight ticket counts among the inspectors is significantly right-skewed. The median number of tickets per inspector is `r py$inspector_spectrum[1,6]`; whereas, the mean is approximately `r round(py$inspector_spectrum[1,2], 0)` tickets, a differencing owing to outlier inspectors with upwards of  `r round(py$inspector_spectrum[1,8], 0)` blight tickets written as shown in `r kfigr::figr(label = "inspector_top10", prefix = TRUE, link = TRUE, type="Table")`. We'll need to determine, during the exploratory phase, the degree to which these outliers influence the regression and decide upon an appropriate treatment.

`r kfigr::figr(label = "inspector_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Inspectors by Blight Tickets Written
```{r inspector_top10}
knitr::kable(as.data.frame(py$inspector_top10, row.names = NULL), digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

No data quality issues appear extant.  That said, we may want to consider adding two new variables to capture the quantity and quality of tickets issued by an inspector in terms of the overall number of tickets written and the percentage of those tickets which were compliant. This will allow us to categorize inspectors based upon compliance percentage, thereby reducing the number categorical levels to something more tractable.  Let's put a pin in it for now and revisit this during the exploratory phase.

* * * 
### Violation Code
```{python violation, code = readLines("../ida.py")[104:113]}
```

`r kfigr::figr(label = "violation_summary", prefix = TRUE, link = TRUE, type="Table")` confirms that each of the `r py$violation_summary$observations` observations was written against one of `r py$violation_summary$unique` violation codes. There are no missing values.

`r kfigr::figr(label = "violation_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Violation Code Summary
```{r violation_summary}
knitr::kable(py$violation_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python violation_plot, code = readLines("../ida.py")[115:117]}
```
`r kfigr::figr(label = "violation_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Violation Code Frequency Histogram

`r kfigr::figr(label = "violation_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Violation Code Frequency Analysis
```{r violation_spectrum}
knitr::kable(py$violation_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Again, we have another significantly right-skewed frequency distribution.  Whereas half of the violation codes have `r py$violation_spectrum[1,6]` or fewer blight tickets, there are others with upwards of `r py$violation_spectrum[1,8]` tickets as shown in `r kfigr::figr(label = "violation_top10", prefix = TRUE, link = TRUE, type="Table")`.

`r kfigr::figr(label = "violation_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Violations by Blight Tickets Written
```{r violation_top10}
knitr::kable(py$violation_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

There are no obvious data quality issues with this variable, and as with inspector, we may also consider the quantity of past tickets written against a violation code as well as the percentage of tickets that were compliant. As with inspector, we can create a new violation code compliance variable based upon compliance percentage thereby reducing the number of categorical levels associated with violations.

Again, we'll take a note and re-address this at feature engineering time.

* * * 
### Violator 
```{python violator, code = readLines("../ida.py")[123:132]}
```

As indicated below, there are `r py$violator_summary$missing` missing values for the violator_name variable. Rather than removing these observations, we can impute the missing violation names with the violation street address.

`r kfigr::figr(label = "violator_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Violator Summary
```{r violator_summary}
knitr::kable(py$violator_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python violator_plot, code = readLines("../ida.py")[134:136]}
```
`r kfigr::figr(label = "violator_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Violator Frequency Histogram

`r kfigr::figr(label = "violator_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Violator Frequency Analysis
```{r violator_spectrum}
knitr::kable(py$violator_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As depicted in `r kfigr::figr(label = "violator_plot", prefix = TRUE, link = TRUE, type="Figure")` and in `r kfigr::figr(label = "violator_spectrum", prefix = TRUE, link = TRUE, type="Table")`, most violators have just a single violation; however, there are repeat offenders with upwards of `r py$violator_spectrum$max` violations.  The top 10 offenders are listed in `r kfigr::figr(label = "violator_top10", prefix = TRUE, link = TRUE, type="Table")`.

`r kfigr::figr(label = "violator_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Violations by Blight Tickets Written
```{r violator_top10}
knitr::kable(py$violator_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Again, we see no obvious data quality issues with this variable; as such, we may consider differentiating repeat offenders by assigning violators to one of several groups, based upon the number of offenses. A compliance rate categorical variable will allow us to categorize violators in terms of past compliance.

* * * 
### Violation Street
```{python violation_street, code = readLines("../ida.py")[142:151]}
```

Each of the `r py$violation_street_summary$observations` observations was written against a parcel on one of `r py$violation_street_summary$unique` streets in Detroit.  There are no missing values for this variable.

`r kfigr::figr(label = "violation_street_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Street Summary
```{r violation_street_summary}
knitr::kable(py$violation_street_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python violation_street_plot, code = readLines("../ida.py")[153:155]}
```
`r kfigr::figr(label = "violation_street_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Street Frequency Histogram

`r kfigr::figr(label = "violation_street_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Street Frequency Analysis
```{r violation_street_spectrum}
knitr::kable(py$violation_street_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As depicted in `r kfigr::figr(label = "violation_street_plot", prefix = TRUE, link = TRUE, type="Figure")` and in `r kfigr::figr(label = "violation_street_spectrum", prefix = TRUE, link = TRUE, type="Table")`, the median number of tickets per street is `r py$violation_street_spectrum[1,6]`; however, the mean is `r round(py$violation_street_spectrum[1,2],0)` tickets per street. This is owing to outlier streets which have upwards of `r py$violation_street_spectrum$max` blight tickets written. We'll need to examine the effects of the outliers during exploration. 

The 10 most frequently tickets streets are listed in `r kfigr::figr(label = "violation_street_top10", prefix = TRUE, link = TRUE, type="Table")`.

`r kfigr::figr(label = "violation_street_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Streets by Blight Tickets Written
```{r violation_street_top10}
knitr::kable(py$violation_street_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Once more, we should consider characterizing streets by the number of tickets written and the percentage of those tickets that were compliant.

* * * 

### City (Mailing)
```{python city, code = readLines("../ida.py")[161:170]}
```

The city variable corresponds to the mailing address of the violator, distinct from the violation address.  There is a city value for each of the `r py$city_summary$observations` observations. 

`r kfigr::figr(label = "city_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Mailing City Summary
```{r city_summary}
knitr::kable(py$city_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python city_plot, code = readLines("../ida.py")[172:174]}
```
`r kfigr::figr(label = "city_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Mailing City Frequency Histogram

`r kfigr::figr(label = "city_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Mailing City Frequency Analysis
```{r city_spectrum}
knitr::kable(py$city_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As depicted in `r kfigr::figr(label = "city_plot", prefix = TRUE, link = TRUE, type="Figure")` and in `r kfigr::figr(label = "city_spectrum", prefix = TRUE, link = TRUE, type="Table")`, the median number of tickets per street is `r py$city_spectrum[1,6]`; however, the mean is `r round(py$city_spectrum[1,2],0)` tickets per city. This is because nearly all violators live in Detroit.

`r kfigr::figr(label = "city_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Mailing Cities by Blight Tickets Written
```{r city_top10}
knitr::kable(py$city_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Any analysis based upon city would require some cleaning and normalization of this field as evidenced by the various spellings and representations of the city of Detroit in `r kfigr::figr(label = "city_top10", prefix = TRUE, link = TRUE, type="Table")`.

* * * 
### State
```{python state, code = readLines("../ida.py")[181:190]}
```

`r kfigr::figr(label = "state_summary", prefix = TRUE, link = TRUE, type="Table")` indicates that there are `r py$state_summary$missing` missing values for state. This can be a consequence of having non-US violators or omission. We can adopt the following imputation strategy. If the country variable is US, we'll assume the state is Michigan.  For NON-US violators, we'll leave the value null as is.

`r kfigr::figr(label = "state_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Mailing State Summary
```{r state_summary}
knitr::kable(py$state_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python state_plot, code = readLines("../ida.py")[192:194]}
```
`r kfigr::figr(label = "state_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Mailing State Frequency Histogram

`r kfigr::figr(label = "state_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Mailing State Frequency Analysis
```{r state_spectrum}
knitr::kable(py$state_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As depicted in `r kfigr::figr(label = "state_plot", prefix = TRUE, link = TRUE, type="Figure")` and in `r kfigr::figr(label = "state_spectrum", prefix = TRUE, link = TRUE, type="Table")`, the median number of tickets per street is `r py$state_spectrum[1,6]`; however, the mean is `r round(py$state_spectrum[1,2],0)` tickets per state. This is because nearly all violators live in Detroit.

`r kfigr::figr(label = "state_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Mailing Cities by Blight Tickets Written
```{r state_top10}
knitr::kable(py$state_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```


* * * 
### Zip Code
```{python zip_code, code = readLines("../ida.py")[200:209]}
```

The zip code variable pertains to the mailing address of the violator, and not the violation address. The training set, as indicated in `r kfigr::figr(label = "zip_code_summary", prefix = TRUE, link = TRUE, type="Table")`, has `r py$zip_code_summary$missing` missing value for zip_code. 

`r kfigr::figr(label = "zip_code_summary", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Zip Code Summary
```{r zip_code_summary}
knitr::kable(py$zip_code_summary, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

```{python zip_code_plot, code = readLines("../ida.py")[211:213]}
```
`r kfigr::figr(label = "zip_code_plot", prefix = TRUE, link = TRUE, type="Figure")`: Blight Ticket by Zip Code Frequency Histogram

`r kfigr::figr(label = "zip_code_spectrum", prefix = TRUE, link = TRUE, type="Table")`: Blight Ticket by Zip Code Frequency Analysis
```{r zip_code_spectrum}
knitr::kable(py$zip_code_spectrum, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As depicted in `r kfigr::figr(label = "zip_code_plot", prefix = TRUE, link = TRUE, type="Figure")` and in `r kfigr::figr(label = "zip_code_spectrum", prefix = TRUE, link = TRUE, type="Table")`, the median number of tickets per zip code is `r py$zip_code_spectrum[1,6]`; however, the mean is `r round(py$zip_code_spectrum[1,2],0)` tickets per zip_code. Another right-skewed distribution with some zip codes having up to `r py$zip_code_spectrum$max` violations as revealed in `r kfigr::figr(label = "zip_code_top10", prefix = TRUE, link = TRUE, type="Table")`.

`r kfigr::figr(label = "zip_code_top10", prefix = TRUE, link = TRUE, type="Table")`: Top 10 Zip Codes by Blight Tickets Written
```{r zip_code_top10}
knitr::kable(py$zip_code_top10, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Aside from the single missing value, there are no obvious data quality issues with this variable. That said, the zip code variable may be too granular to be useful as a predictor since it would require `r py$zip_code_summary$unique` categorical levels.  We may consider creating a new variable from the first three numbers of the zip code which pertain to regional areas.

* * * 
### Country
```{python country, code = readLines("../ida.py")[219:225]}
```
Blight violators hale from one of 5 countries; however nearly 100% are US based.

`r kfigr::figr(label = "country_tbl", prefix = TRUE, link = TRUE, type="Table")`: Blight Tickets by Country
```{r country_tbl}
knitr::kable(py$country, digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

Given this inbalance, country would not be a useful categorical variable for analysis and prediction.

* * * 
### Latitude / Longitude
```{python lat_lon, code = readLines("../ida.py")[231:234]}
```

The following rather course summary of latitude and longtitude shows no glaring data quality problems aside from the two missing variables.

`r kfigr::figr(label = "ida_lat_lon", prefix = TRUE, link = TRUE, type="Table")`: Summary of Latitude and Longitude Variables
```{r ida_lat_lon}
knitr::kable(py$lat_lon, digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

This data will be useful for a geolocation based visualization of the blight ticket data.

* * * 
### Dates
```{python dates, code = readLines("../ida.py")[241:252]}
```
Ticket issued dates appear to be in satisfactory condition and there are no missing dates; however we do have tickets dated back to 1988 in the training set. According to the data source, the ticket dates were to start in 2005.  We'll need to examine such outliers during the next phase. 

Two additional issues arise upon examination of the hearing dates. First, there are over 200 missing hearing dates and there are `r nrow(py$errors)` observations where the hearing date did not follow the date in which the ticket was issued. A sample of the potential errors are listed in `r kfigr::figr(label = "date_errors", prefix = TRUE, link = TRUE, type="Table")`.

`r kfigr::figr(label = "dates_summary", prefix = TRUE, link = TRUE, type="Table")`: Summary of Ticket Issued and Hearing Dates
```{r dates_summary}
knitr::kable(py$dates_summary, digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

`r kfigr::figr(label = "date_errors", prefix = TRUE, link = TRUE, type="Table")`: Sampling of Date Errors
```{r date_errors}
knitr::kable(py$sample_errors, digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

For both the missing and presumably incorrect hearing dates, new hearing dates may be imputed based upon the mean interval between ticket issued dates and hearing dates, overall or at the violation code level. We should also create a variable to capture the interval between the ticket issued date and the hearing date. A variable that represents the financial costs vis-a-vis the time allotted for payment might be interesting as well.  

* * * 
### Judgment Amount
```{python ja, code = readLines("../ida.py")[258:261]}
```
Alas, we come to the last variable for this univariate exploration. Judgment amounts covered a significant range from \$0 upwards to \$`r round(py$ja_distribution$max, 0)`. An inspection of the data revealed `r nrow(py$zero_ja)` observations with zero judgment amounts. All of these blight tickets were judged 'compliant by no fine'. 


```{python ja_plot, code = readLines("../ida.py")[263:265]}
```
`r kfigr::figr(label = "ja_plot", prefix = TRUE, link = TRUE, type="Figure")`: Distribution of Judgment Amounts

`r kfigr::figr(label = "ja_distribution", prefix = TRUE, link = TRUE, type="Table")`: Distribution of Judgment Amounts
```{r ja_distribution}
knitr::kable(py$ja_distribution, digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "center") 
```

As indicated in `r kfigr::figr(label = "ja_distribution", prefix = TRUE, link = TRUE, type="Table")`, both the mean and the mode judgment amount is \$`r round(py$ja_distribution[1,6], 0)`. There were no obvious date quality issues with this variable; however, outliers must be examined in terms of their influence on regression.

* * * 

### Summary
To wrap up this part of the analysis, let's quickly summarize our actions, observations and implications for next steps. We've reduced the data set from `r nrow(py$df)` blight ticket observations and `r ncol(py$df)` variables to `r nrow(py$df)`  observations and `r ncol(py$df)` variables. Aside from a few missing values for violator name, state, zip code, and hearing date and spelling variation in the city variable, there were no glaring data quality issues.  However, the frequency distributions of blight tickets all categorical variables were significantly right skewed. A thorough outlier assessment should be conducted during the exploration phase to ascertain the influence of such outliers on the regression lines.

Several data preparation treatments were identified. Concretely:   
**Agency**   
* Combine the agency variable categorical levels for Neighborhood City Halls and Detroit Police Department into a single categorical level.  
* Create the agency variable to capture the number of past blight tickets.  
* Create the agency variable, which is the percentage of past tickets which were compliant.   


**Inspector**    
* Create the inspector_tickets variable to capture the number of past blight tickets for each inspector ticket.   
* Create the inspector_compliance variable, which is the percentage of past tickets which were compliant.  

**Violation Code**   
* Create the violation_code_tickets variable to capture the number of past blight tickets for each violation_code.   
* Create the violation_code_compliance variable, which is the percentage of past tickets which were compliant.  

**Violator**    
* Create the violator_tickets variable to capture the number of past blight tickets written for a violator.   
* Create the violator_compliance variable, which is the percentage of past tickets which were compliant.  

**Violation Street**   
* Create the violation_street_tickets variable to capture the number of past blight tickets for each violation_street.
* Create the violation_street_compliance variable, which is the percentage of past tickets which were compliant.  

**City**  
* Create a new binary categorical variable that indicates whether the mailing city for the owner is out of town.  
* Standardize spelling of Detroit. 

**State**  
* Create a new binary categorical variable that indicates whether the mailing city for the owner is out of state.  
* Create the state variable to capture the number of past blight tickets by owner state of residence.   
* Create the state variable, which is the percentage of past tickets which were compliant.   
  
**Region**   
* Create a region variable based upon the first three numbers of the zip code.   
* Create the region_tickets variable to capture the number of past blight tickets for each region.   
* Create the region_compliance variable, which is the percentage of past tickets which were compliant.  

**Dates**   
* Create a payment_window variable which captures the number of days between the ticket issued date and the hearing date.  
* Hearing dates that occur on or before the ticket issued dates will be imputed based upon the median payment window.  
* Missing hearing dates will be imputed based upon the median payment_window.  


**Judgment Amount**
* Create a new variable representing the log (x+1) transformation of the judgment amount. 
 

This concludes the initial data analysis phase. Next, the exploratory data analysis phase will address the primary research question: which variables best predict compliance. 